Supporting Non-Functional Requirements in Services Software Development Process: An MDD Approach

### 2013-10-21 SOFSEM2014/FinalVersion/main.tex
Update main.tex

Despite the variety of techniques proposed, there is not yet a consensus on a software
process methodology for web services. Some methodologies address the service-based 
development towards a standard or a new way to develop reliable applications. 
SOD-M and SOMF \cite{somf} are MDD approaches for web services; 
S-Cube \cite{scube2010book} is focused on the representation of business processes and 
service-based development; SOMA \cite{soma} is a methodology described by IBM
for SOA solutions; Extended SOA \cite{PapazoglouH06} merges RUP and
BPM\cite{bpm} concepts for service modeling; DEVISE \cite{DEVISE} is a
methodology for building service-based infrastructure for collaborative
enterprises. Other proposals include, the WIED model \cite{TongrungrojanaL04}, 
that acts as a bridge between business modeling and design models, and traditional 
approaches for software engineering \cite{sommerville08} applied to SOC.

>>>>

More specifics works dealing with non-functional properties in service-oriented 
development can be organized in two main groups: those working on the \textit{modeling 
of particular non-functional properties or QoS attributes} and those 
proposing \textit{architectures or frameworks to manage and 
validate QoS attributes in web service composition processes}. 
First type of works are thus considering a particular non-functional concern (e.g. security)
which is modeled and then "glued" or associated to functional models of an application. 
The work of Chollet et al.~\cite{CholletL09} defines a proposal to associate non-functional 
quality properties (security properties in that case) to functional activities which are modeled 
in a web service composition model. Schmeling et al.~\cite{SchmelingCM11} present 
an approach and also a toolset for specifying and implementing non-functional 
concerns in web service compositions; non-functional concerns are modeled first 
in non-functional model and then related to a service composition 
represented in a BPMN diagram. Ovaska et al.~\cite{OvaskaEHPA10} present an approach to 
support quality management at design time. Quality requirements are modeled in a first phase 
and represented then in an architectural model where quality requirements are associated to some
components in the model. In our case we propose to model non-functional properties joint 
to the modeling of functional concerns what simplifies the web service development process. 
Non-functional properties are represented in our work as constraints and policies 
but more general quality model such as the proposed by ~\cite{Goeb2011,Klass2009} 
could be taken in consideration in further works. Second type of works related to 
our proposal proposes specifics architecture or frameworks to manage and validate 
QoS attributes in service composition processes ~\cite{XiaoCZBOLH08,Babamir2010,Karunamurthy2012787}. 
Although our proposal is focusing on the design of service oriented application with non-functional 
properties and not on the validation of QoS attributes at runtime.

Despite the variety of techniques proposed, there is not yet a consensus on a software methodology 
for web services. Some methodologies address the service-based 
development towards a standard or a new way to develop reliable applications. 
SOD-M and SOMF \cite{somf} are MDD approaches for web services; 
S-Cube \cite{scube2010book} is focused on the representation of business processes and 
service-based development; SOMA \cite{soma} is a methodology described by IBM
for SOA solutions; Extended SOA \cite{PapazoglouH06} merges RUP and
BPM\cite{bpm} concepts for service modeling; DEVISE \cite{DEVISE} is a
methodology for building service-based infrastructure for collaborative
enterprises. Other proposals include, the WIED model \cite{TongrungrojanaL04}, 
that acts as a bridge between business modeling and design models, and traditional 
approaches for software engineering \cite{sommerville08} applied to SOC.
